---
title: "Data Wrangling and Analysis in R"
output: html_notebook
---


## 1. Installing packages 
In R, packages are a collection of functions that provide functinalities not given by the R core functionalities. We'll install tidyverse packages which will provide us with packages for performing data manipulation and analysis.

#### Let's install tidyverse packages in Rstudio
```{r}
install.packages("tidyverse")
```

#### Load tidyverse for use in our notebook
```{r}
library(tidyverse)
```

## 2. Data Importation

```{r}
# a. Importing from a CSV file from URL and loading the entire dataframe.

MTN_df <- read_csv("https://bit.ly/2ZlpzjF")
MTN_df
```
## 3. Data Understanding


```{r}
# We can preview a specific number of records by passing a second argument to the head() function:

head(MTN_df, 10)
```

```{r}
# To preview the last 6 records of a dataset, we use the tail function, padding the dataframe name as the only argument.

tail(MTN_df)
```

```{r}
# The `glimpse()` function displays the variable names, datatypes and first few records of the dataset.

glimpse(MTN_df)

```
```{r}
# In addition to what was printed using the glimpse function, if we want to display more information about the datatypes we can use the `str()` function and pass the dataframe name as the only argument.

str(MTN_df)
```


```{r}
# Sometimes we might want to preview an unbiased number of dataset records. We use the sample_n() function and pass the dataframe name and number of samples as the arguments. 

sample_n(MTN_df, 10)
```

```{r}
# Lastly, to quickly check the size of the dataset without displaying additional information, we use the `dim ()` function, passing the dataframe name as the only argument.

dim(MTN_df)
```

## 4. Performing Standardisation

```{r}
# Get column names of the data dataset
names(MTN_df)
```



```{r}
# Strip the leading and trailing spaces
names(MTN_df) <- trimws(names(MTN_df), which = c("both"))
names(MTN_df)
```

```{r}
# As a standard while working with colunn names we usually use lowercase characters. Let's transform all column names to have lowercase characters using the `tolower()` function. We pass a vector containing the column names to the function.
names(MTN_df) <- tolower(names(MTN_df))
names(MTN_df)
```

## 5. Handling Irrelevant Data

```{r}
# Let's drop the monthlycharges and totalcharges columns

MTN_df = select(MTN_df, -c("monthlycharges", "totalcharges"))
head(MTN_df)
```



## 6. Finding and dealing with Duplicates 

Sometimes duplicates can mislead our analysis. Let's find and then leave the out from our data analysis dataset.

```{r}
# Let's determine the number of records using dim()
dim(MTN_df)
```

```{r}
# To extract the duplicated records from the dataset, we use x[duplicated(x),], where x is the dataframe name.

MTN_df[duplicated(MTN_df), ]
```

```{r}
# If we don't want to extract records which are not duplicates we add a logical negation `!` infront of the duplicated function.
MTN_df[!duplicated(MTN_df), ]


```


## 7. Missing data

To avoid wrong data analysis conclusions, we find and deal with missing data.


```{r}
# We use the `is.na()` function to identify missing data and pass the dataframe. The output given has TRUE or FALSE to represent whether the value is missing or not.
is.na(MTN_df)

```

```{r}
# To find the total missing values in each column we use the `colSums()` function and pass is.na(dt) as the argument and where `dt` is the dataframe.

colSums(is.na(MTN_df))
```

# Drop all rows containing missing values.
```{r}
# We can drop all records containing missing values using the na.omit() function and passing the dataframe name as the argument.
na.omit(MTN_df)

```
## QUESTION 1. total number of retained and churned customers

```{r}
## total number of churned customers
#filter(MTN_df, churn =='Yes')
count(MTN_df, churn)
```



## QUESTION 2. average tenure:how long they have been customers
```{r}
MTN_df %>%
  group_by(gender, churn) %>%
  summarize(Average_tenure = mean(tenure))

```


## QUESTION 3. demographics-age and gender
```{r}
##Churn by age and gender
MTN_df %>%
  filter (churn == 'Yes')%>%
  group_by(gender,seniorcitizen) %>%
  summarize(Total_Churn = length(churn))
```




